---
title: "GeoCoded Data Exploration"
author: "Matty Samach"
date: "12/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup}
library(dplyr)
library(ggplot2)
#setwd("Dropbox/Documents/CMU/Classes/Fall 2019/Applied ML/Project/Data/GeoCoded/")

# Get list of unique restaurant addresses from inspection data
insp_restaurants = read.delim("Inspections_Geocoded.txt", stringsAsFactors = F) %>% select(Match_addr, USER_id)
insp_restaurants = insp_restaurants %>% rename(restaurant = USER_id) %>% distinct()
```

```{r Inspection restaurants}
insp_groups = insp_restaurants %>% group_by(Match_addr) %>% 
  summarise(num_restaurants = n())
  
insp_groups %>% arrange(desc(num_restaurants))
```
There are 529 rows that don't have an address at all. Will exclude this one and explore the distribution more deeply.

```{r}
insp_summaries = insp_groups %>% filter(Match_addr != "") %>% 
  group_by(num_restaurants) %>% summarize(count=n())

insp_summaries %>% arrange(count)
```

Within the inspectiond ataset only, there are 5274 addresses that correspond to only one restaurant. Will now extract only these locations.

```{r extract unique inspections}
insp_unique_addr = insp_groups %>% filter(num_restaurants == 1) %>% select(Match_addr)
```

Will now bring in the ReferenceUSA Business data and merge it to these unique addresses. Will then have to check again for multiple addresses per business and filter them out.

```{r RefUSABusinesses}
refUSA = read.delim("Businesses_Geocoded.txt", stringsAsFactors = F) %>% select(Match_addr, USER_Business_ID)
refUSA = refUSA %>% rename(bus_id = USER_Business_ID) %>% distinct()
```

```{r}
joined = insp_unique_addr %>% left_join(refUSA)

joined_grouped = joined %>% group_by(Match_addr) %>% 
  summarize(num_restaurants = n()) %>% arrange(desc(num_restaurants))
```

```{r}
joined_summary = joined_grouped %>% group_by(num_restaurants) %>% 
  summarize(count = n())
```

After left joining inspection restaurants to ReferenceUSA businesses, there are 4199 addresses that only correspond to one restaurant. Will only take these addresses now.

```{r}
unq_rstrnts = joined_grouped %>% filter(num_restaurants == 1) %>% select(Match_addr)
```

Now will join this list of total unique restaurants back to referenceUSA data to check how many of them do actually correspond to a refUSA data point

```{r}
usable_rows = unq_rstrnts %>% inner_join(refUSA) %>% inner_join(insp_restaurants)

nrow(usable_rows)

write.csv(x = usable_rows, file = "usable_addresses.csv", row.names = F)
```

There are 1992 total restaurants that include all refUSA data


